---
title       : Meet R
description : Soft introduction to R

--- type:NormalExercise lang:r xp:50 skills:1

## Basic tools

Welcome. On your right you see the R editor area and below that the R console.  

The way to use these two is to first write commands to the editor area (just as you would write in any text editor). 
Then you can execute the row where your cursor is, or the rows you have highlighted by pressing `Ctrl + Enter`.  

Output will then appear in the console. 

Try it!
  
*** =instructions
- Follow the instructions in the sript and use R as a calculator
- `+`, `-`, `/`, `*`  are the plus, minus, division and multiplication operations.
- When you are done experimenting, press the 'Submit answer' button.

*** =hint
- Remember that Ctrl+Enter executes a single row.


*** =pre_exercise_code
```{r}
```


*** =sample_code
```{r}
# This is the R editor!


# You can use R as a calculator. Execute the below line with Ctrl + Enter.
5 + 5

# A hashtag at the beginning of the line defines the line as a comment

# Executing a comment will only print the comment, try it!
# 5 + 5


# Try your own calculations below, and execute them



# Below is the R console, where you will see output

```

*** =solution
```{r}
# I am a comment telling you that this is the R editor!


# You can use R as a calculator. Execute the below line with Ctrl + Enter
5 + 5


# Executing a comment will only print the comment, try it!
# 5 + 5


# Try your own calculations below, and execute them



# Below is the R console, where you will see output

```

*** =sct
```{r}

# It's always smart to include the following line of code at the end of your SCTs
# It will check whether executing the student's code resulted in an error, 
# and if so, will cause the exercise to fail
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```


--- type:NormalExercise lang:r xp:50 skills:1
## Objects

In R you can create and operate on things called *objects*. 
An object is just something that knows it's own type and can store information such as numerical values and names.

The beauty of R is that you can store these objects for later use  using the assign operator: `<-`.


*** =instructions
- Execute the lines that define the objects `hello` and `i_am_numeric` and print their values
- Assign your name or nickname to `my_name`
- Assign the text "you're awsome!" to `my_message`

*** =hint
- Remember that Ctrl+Enter executes a single row.

*** =pre_exercise_code
```{r}
```


*** =sample_code
```{r}

# Create an object and save it to memory
hello <- "Hi there!"

# Print the object by typing it's name
hello

# Create a numeric object
i_am_numeric <- 5 + 5

# Do further calculations with the object
(i_am_numeric/5) + 2

# You can override the value of hello by assigning a new value to it
hello <- i_am_numeric
hello


# Create my_name and my_message here



```

*** =solution
```{r}

# Create an object and save it to memory
hello <- "Hi there!"

# Print the object by typing it's name
hello

# Create a numeric object
i_am_numeric <- 5 + 5

# Do further calculations with the object
(i_am_numeric/5) + 2

# You can override the value of hello by assigning a new value to it
hello <- i_am_numeric
hello


# Create my_name and my_message here.
my_name <- "Tuomo"
my_message <- "you are awsome!"


```

*** =sct
```{r}

test_object("my_message")

# It's always smart to include the following line of code at the end of your SCTs
# It will check whether executing the student's code resulted in an error, 
# and if so, will cause the exercise to fail
test_error()

# Final message the student will see upon completing the exercise
success_msg(paste0(my_name, ", ", my_message))

```

--- type:NormalExercise lang:r xp:100 skills:1
## Data frames

The most common object to operate on is the `data frame` object, which stores data in tabular form.

In R, the basic idea is to use *functions* to operate on objects. Here you will get to apply some functions that are helpful for inspecting data: `head()` and `summary()`.

You will also learn how to access the columns of a data frame using the `$` sign.

You can use an R data frame called `iris`, which contains information about different flower species. You have access to iris by it's name, `iris`.

*** =instructions
- Use `summary()` to get a summary of the variable Petal.Length in `iris`

*** =hint
- Remember that Ctrl + Enter executes a single row
- The correct answer is of the form `function(iris$column_name)`


*** =pre_exercise_code
```{r}
```


*** =sample_code
```{r}

# Functions take arguments inside the parenthesis. Take a peek of iris
head(iris)

# You can use the $ sign to access columns.
mean(iris$Sepal.Length)


# Create a summary of Petal.Width



```

*** =solution
```{r}

# Functions take objects as arguments inside the parenthesis. Take a peek of iris.
head(iris)

# You can use the $-sign to access columns
mean(iris$Sepal.Length)


# Create a summary of Petal.Width



```

*** =sct
```{r}

test_function("summary", args="object")

# It's always smart to include the following line of code at the end of your SCTs
# It will check whether executing the student's code resulted in an error, 
# and if so, will cause the exercise to fail
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```


--- type:NormalExercise lang:r xp:100 skills:1
## Functions


As we just saw, in R you can operate on objects using *functions*. This is how you're actually going to do all the magic.

During this course you will operate on ready-made functions, but for clarity we will look at some of the details of functions.

A function is called by typing it's name and giving it the necessary *arguments* inside parenthesis.


*** =instructions
- Try the functions that are already defined
- Use `do_maths()` with `n` equal to 20 and `k` equal to 10.
- Create object 'ans' by assigning the above requested value to it.

*** =hint
- Remember that the editor code is executed line by line with Ctrl + Enter
- Assignment is done with the `<-` operator: `object <- function(arguments)`

*** =pre_exercise_code
```{r}
# code to preload datasets etc

add_5 <- function(n) {
  n + 5
}
do_maths <- function(n, k = 5) {
n * k - n + k / k
  }
```


*** =sample_code
```{r}

# This code creates a simple function with one argument: n
add_5 <- function(n) {
  n + 5
}

# Here we call the add_5 function. (It is already defined and saved for you)
add_5(n = 5)
add_5(n = -5)

# Functions can be more complicated and have multiple arguments. Arguments can also have default values
do_maths <- function(n, k = 5) {
n * k - n + k / k
  }

# Here we call the do_maths function. k equals 5 by default so these two lines are equal
do_maths(n = 10, k = 5)
do_maths(n = 10)

# If you give the arguments in the right order, you don't have to state which is which.
do_maths(10)

# So these are the same thing
do_maths(8, 3)
do_maths(k = 3, n = 8)

# Create the object 'ans' here.




```

*** =solution
```{r}

# This code creates a simple function with one argument: n
add_5 <- function(n) {
  n + 5
}

# Here we call the add_5 function. (It is already defined and saved for you)
add_5(n = 5)

# Functions can be more complicated and have multiple arguments. Arguments can also have default values
do_maths <- function(n, k = 5) {
n * k - n + k / k
  }

# Here we call the do_maths function. k = 5 by default so the two lines are equal
do_maths(n = 10, k = 5)
do_maths(n = 10)

# If you give the arguments in the right order, you don't have to state which is which.
do_maths(10)

# So these are the same thing
do_maths(8, 3)
do_maths(k = 3, n = 8)

# Create the object 'ans' here.

ans <- do_maths(20, 10 )



```

*** =sct
```{r}

# tests. 

# test_object("name") 
# test_function("name", args=c("arg1","arg2"))

test_object("ans")

# It's always smart to include the following line of code at the end of your SCTs
# It will check whether executing the student's code resulted in an error, 
# and if so, will cause the exercise to fail
test_error()

# Final message the student will see upon completing the exercise
success_msg("Great job!")

```

--- type:NormalExercise lang:r xp:100 skills:1
## Vectors and more data frames

info about vectors

*** =instructions
- Use `?quantile` to see the help page on `quantile`
- Use an online resource to find the connection of percentiles and quantiles
- Compute the 25th, 50th and 75th percentiles of Sepal.Length.

*** =hint
-  you need two give 2 arguments to `quantile()`


*** =pre_exercise_code
```{r}
```


*** =sample_code
```{r}



# Get the 25th, 50th and 75th percentile of Sepal.Width



```

*** =solution
```{r}


# Get the 25th, 50th and 75th percentile of Sepal.Width
quantile(iris$Sepal.Length, probs = c(0.25, 0.5, 0.75))




```

*** =sct
```{r}


# It's always smart to include the following line of code at the end of your SCTs
# It will check whether executing the student's code resulted in an error, 
# and if so, will cause the exercise to fail
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```