---
title       : Meet R
description : Soft introduction to R

--- type:NormalExercise lang:r xp:50 skills:1

## Basic tools

Welcome. On your right you see the R editor area and below that the R console.  

The way to use these two is to first write commands to the editor area (just as you would write in any text editor). 
Then you can execute the row where your cursor is, or the rows you have highlighted by pressing `Ctrl + Enter`.  

Output will then appear in the console. 

Let's try it!
  
*** =instructions
- Follow the instructions in the sript and use R as a calculator
- `+`, `-`, `/`, `*`  are the plus, minus, division and multiplication operations.
- When you are done experimenting, press the 'Submit answer' button.

*** =hint
- Remember that Ctrl+Enter executes a single row.


*** =pre_exercise_code
```{r}
```


*** =sample_code
```{r}
# This is the R editor!


# You can use R as a calculator. Execute the below line with Ctrl + Enter.
5 + 5

# A hashtag at the beginning of the line defines the line as a comment.

# Executing a comment will only print the comment, try it!
# 5 + 5


# Try your own calculations below, and execute them.



# Below me is the R console, where you will see output

```

*** =solution
```{r}
# I am a comment telling you that this is the R editor!


# You can use R as a calculator. Execute the below line with Ctrl + Enter
5 + 5


# Executing a comment will only print the comment, try it!
# 5 + 5


# Try your own calculations below, and execute them.



# Below me is the R console, where you will see output

```

*** =sct
```{r}

# It's always smart to include the following line of code at the end of your SCTs
# It will check whether executing the student's code resulted in an error, 
# and if so, will cause the exercise to fail
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```


--- type:NormalExercise lang:r xp:100 skills:1
## R objects

In R you can create and operate on things called *objects*. 
An object is just something that knows it's own type and can store information such as numerical values and names.

The beauty of R is that you can store these objects for later use in your very own variables using the assign operator: `<-`.


*** =instructions
- Execute the lines that define the variables `hello` and `ans` and print their values
- Create your own variable, called `my_name` and assign your name or nickname to it.
- Create your own variable, called `my_message` and assign the text "you are awsome!" to it.

*** =hint
- Remember that Ctrl+Enter executes a single row.

*** =pre_exercise_code
```{r}
```


*** =sample_code
```{r}

# Create a variable named hello.
hello <- "Hi there!"

# Print the variable by typing it's name.
hello

# Create a numeric variable.
ans <- 5 + 5

# Do some operations with ans.
(ans/5) + 2


# Create my_name and my_message here.



```

*** =solution
```{r}

# Create a variable named hello.
hello <- "Hi there!"

# Print the variable by typing it's name.
hello

# Create a numeric variable.
ans <- 5 + 5

# Do some operations with ans.
(ans/5) + 2


# Create my_name and my_message here
my_name <- "Tuomo"
my_message <- "you are awsome!"


```

*** =sct
```{r}

test_object("my_message")

# It's always smart to include the following line of code at the end of your SCTs
# It will check whether executing the student's code resulted in an error, 
# and if so, will cause the exercise to fail
test_error()

# Final message the student will see upon completing the exercise
success_msg(paste0(my_name, ", ", my_message))

```


--- type:NormalExercise lang:r xp:100 skills:1
## R functions


Now that we know how to create variables that store our objects, the real fun can begin!

In R you can operate on objects using *functions*. This is how we're actually going to do all the magic. During this course we will operate on ready-made functions, but for clarity we will look at how a function is created and what are it's parts.

A function is called by typing it's name and giving it the necessary *arguments* inside parenthesis.


*** =instructions
- Try the function calls that are already defined
- use the `create_nick()`- function to create a nick with the name "R" and number 93.
- assign the created nick to a variable called 'ans'

*** =hint
- Remember that the editor code is executed line by line with Ctrl + Enter
- Assignment is done with the `<-` operator: `variable_name <- function_name(arguments)

*** =pre_exercise_code
```{r}
# code to preload datasets etc
add_5 <- function(n) {
  result <- n + 5
  result
}

create_nick <- function(name = "Juha", no = 86) {
  paste("Your nick is: ", name, no, sep="")
}
```


*** =sample_code
```{r}

# This code creates a simple function with one argument: n
add_5 <- function(n) {
  result <- n + 5
  result
}

# Here we call the add_5 function. (It is already defined and saved for you)
add_5(n = 5)

# Functions can be more complicated and have multiple arguments. Arguments can also have default values
create_nick <- function(name, no = 86) {
  paste("Your nick is: ", name, no, sep="")
}

# Here we call the create_nick function
create_nick(name="Juha")

# If you give the arguments in the right order, you don't have to state which is which.
create_nick("Juha")
create_nick("Minni", 98)
create_nick(no = 98, name = "Minni")


# Create the variable 'ans' here.




```

*** =solution
```{r}


# Here we call the add_5 function. (It is already defined and saved for you)
add_5(n = 5)

# Here we call the create_nick function
create_nick(name="Juha")

# If you give the arguments in the right order, you don't have to state which is which.
create_nick("Juha")
create_nick("Minni", 98)
create_nick(no = 98, name = "Minni")


# Assign the function output to a variable 'ans' here

ans <- create_nick("R", 93)



```

*** =sct
```{r}

# tests. 

# test_object("name") 

# test_function("name", args=c("arg1","arg2"))

test_object("ans")

# It's always smart to include the following line of code at the end of your SCTs
# It will check whether executing the student's code resulted in an error, 
# and if so, will cause the exercise to fail
test_error()

# Final message the student will see upon completing the exercise
success_msg("Great job!")

```


--- type:NormalExercise lang:r xp:100 skills:1
## R data frames

The most common object to operate on is the `data frame` object, which stores data in tabular form.

Here we will look at some of the functions that are helpful for inspecting a data frame: `str()`, `names()`, `head()`, `tail()`,  We will use an R data frame called `iris`, which contains information about different flower species.

You have access to iris by it's name, `iris`.

*** =instructions
- Use the `names()` function to see which variables are in the iris data frame
- Use the `str()` function to look at the structure of `iris` and the type of it's variables
- Use the the command `?head` to see the help page on head. Then use `head()` to see the first 5 observations of iris.
- Guess what the `tail()` function does?

*** =hint
- Remember that Ctrl+Enter executes a single row.
- All the functions take the iris data frame object as an argument. So you should type ´iris` inside the function parenthesis.
- The `x` argument of `head(x, n)` defines the data frame and the `n` argument defines the number of observations.


*** =pre_exercise_code
```{r}
```


*** =sample_code
```{r}

# Look at the variables of iris


# Look at the structure of iris


# Look at the top 5 rows of iris



```

*** =solution
```{r}

# Look at the variables of iris
names(iris)

# Look at the structure of iris
str(iris)

# Look at the top 5 rows of iris
head(iris, n = 5)


```

*** =sct
```{r}

test_function("names",args="x")
test_function("str",args="x")
test_function("head", args=c("x","n"))

# It's always smart to include the following line of code at the end of your SCTs
# It will check whether executing the student's code resulted in an error, 
# and if so, will cause the exercise to fail
test_error()

# Final message the student will see upon completing the exercise
success_msg(paste0(myName,"_",myBD,", ", myMessage))

```
