---
title       : Meet R
description : Soft introduction to R

--- type:NormalExercise lang:r xp:50 skills:1
## Basic tools

Welcome. On your right you see the R editor area and below that the R console. 
The way to use these two is to first write commands to the editor area (just as you would write in any text editor). 
Then you can execute the row where your cursor is, or the rows you have highlighted by pressing Ctrl+Enter. 
Output will then appear in the console. 

Let's try it!
  
*** =instructions
- Follow the instructions in the sript and use R as a calculator
- `+`, `-`, `/`, `*`  are the plus, minus, division and multiplication operations.
- When you are done experimenting, press the 'Submit answer' button.

*** =hint
- Remember that Ctrl+Enter executes a single row.
  
  *** =pre_exercise_code
```

*** =sample_code
```{r}
# I am a comment telling you that this is the R editor!


# You can use R as a calculator. Execute the below line with Ctrl + Enter
5 + 5


# Executing a comment will only print the comment, try it!
# 5 + 5


# Try your own calculations below, and execute them.



# Below me is the R console, where you will see output
```

*** =solution
```{r}
# I am a comment telling you that this is the R editor!


# You can use R as a calculator. Execute the below line with Ctrl + Enter
5 + 5


# Executing a comment will only print the comment, try it!
# 5 + 5


# Try your own calculations below, and execute them.



# Below me is the R console, where you will see output
```

*** =sct
```{r}

# It's always smart to include the following line of code at the end of your SCTs
# It will check whether executing the student's code resulted in an error, 
# and if so, will cause the exercise to fail
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1
## R objects

In R you can create and operate on things called *objects*. 
An object is just something that knows what kind of object it is and can store information. 
The beauty of R is that you can store these objects for later use in your very own variables using the assign operator: `<-`.


Let's try it!

*** =instructions
- Execute the lines that define the variables `myHello` and `ans` and print their values
- Create your own variable, called `myMessage` and assign the text "I am an awsome R message." to it.

*** =hint
- Remember that Ctrl+Enter executes a single row.

*** =pre_exercise_code
```

*** =sample_code
```{r}

# create a variable named myHello
myHello <- "Hi there!"

# print the variable by simply typing it's name
myHello

# Create a numeric variable
ans <- 5 + 5

# Do some operations with ans
(ans/5) + 2


# Create myName and myMessage here



```

*** =solution
```{r}

# create a variable named myHello
myHello <- "Hi there!"

# print the variable by simply typing it's name
myHello

# Create a numeric variable
ans <- 5 + 5

# Do some operations with ans
(ans/5) + 2


# Create myName and myMessage here
myName <- "Tuomo"
myMessage <- "you're an awsome R coder!"


```

*** =sct
```{r}

test_object("myMessage")

# It's always smart to include the following line of code at the end of your SCTs
# It will check whether executing the student's code resulted in an error, 
# and if so, will cause the exercise to fail
test_error()

# Final message the student will see upon completing the exercise
success_msg(paste(myName, myMessage))

```
