---
title       : Meet R
description : Soft introduction to R

--- type:NormalExercise lang:r xp:100 skills:1

## Basic tools

Welcome! On your right you see the R editor area and below that the R console.  

The way to use these two is to first write commands to the editor area (just as you would write in any text editor). 
Then you can tell R that you want to execute a command by pressing `Ctrl + Enter`. This executes the row where your cursor currently is.

Input and output will then appear in the console. 

Try it!
  
*** =instructions
- Follow the instructions in the sript and use R as a calculator
- `+`, `-`, `/`, `*`  are the plus, minus, division and multiplication operations.
- When you are done experimenting, press the 'Submit answer' button.

*** =hint
- Remember that Ctrl+Enter executes a single row.


*** =pre_exercise_code
```{r}
```


*** =sample_code
```{r}
# This is the R editor!


# You can use R as a calculator. Execute the below line with Ctrl + Enter.
5 + 5

# A hashtag at the beginning of the line defines the line as a comment

# Executing a comment will only print the comment, try it!
# 5 + 5


# Try your own calculations below, and execute them



# Below is the R console, where you will see output

```

*** =solution
```{r}

```

*** =sct
```{r}

# It's always smart to include the following line of code at the end of your SCTs
# It will check whether executing the student's code resulted in an error, 
# and if so, will cause the exercise to fail
test_error()

test_output_contains("10", incorrect_msg="Please execute the code `5 + 5`")


# Final message the student will see upon completing the exercise
success_msg("Good work!")

```


--- type:NormalExercise lang:r xp:100 skills:1
## Objects

In R you can create and operate on things called *objects*. 
An object is just something that knows it's own type and can store information such as numerical values and names.

The beauty of R is that you can store these objects for later use  using the assign operator: `<-`.


*** =instructions
- Execute the lines that define the objects `hello` and `i_am_numeric` and print their values
- Assign your name or nickname to `my_name`
- Assign the text "you're awsome!" to `my_message`

*** =hint
- Remember that Ctrl+Enter executes a single row.

*** =pre_exercise_code
```{r}
```


*** =sample_code
```{r}

# Create an object and save it to memory
hello <- "Hi there!"

# Print the object by typing it's name
hello

# Create a numeric object
i_am_numeric <- 5 + 5

# Do further calculations with the object
(i_am_numeric/5) + 2

# You can override the value of hello by assigning a new value to it
hello <- i_am_numeric
hello


# Create my_name and my_message here



```

*** =solution
```{r}

# Create an object and save it to memory
hello <- "Hi there!"

# Print the object by typing it's name
hello

# Create a numeric object
i_am_numeric <- 5 + 5

# Do further calculations with the object
(i_am_numeric/5) + 2

# You can override the value of hello by assigning a new value to it
hello <- i_am_numeric
hello


# Create my_name and my_message here.
my_name <- "Tuomo"
my_message <- "you are awsome!"


```

*** =sct
```{r}

test_object("hello")
test_output_contains("hello")
test_object("my_message")

test_error()

success_msg(paste0(my_name, ", ", my_message))

```

--- type:NormalExercise lang:r xp:150 skills:1,3
## Data frames

The most common object to operate on is the `data frame` object, which stores data in tabular form.

In R, the basic idea is to use *functions* to operate on objects. Here you will get to apply some functions that are helpful for inspecting data: `head()` and `summary()`.

You will also learn how to access the columns of a data frame using the `$`-sign.

You can use an R data frame called `iris`, which contains information about different flower species. You have access to iris by it's name, `iris`.

*** =instructions
- Execute the functions `head()` and `mean()` and see the help page of `mean()`
- Use `summary()` and `$` to create a summary of `Petal.Width` in `iris`

*** =hint
- Remember that Ctrl + Enter executes a single row
- The correct answer is of the form `function(iris$column)`


*** =pre_exercise_code
```{r}
```


*** =sample_code
```{r}

# Functions take arguments inside the parenthesis. Take a peek of iris.
head(iris)

# You can use ?function or help(function) to open a help page.
?mean

# You can use the $-sign to access columns of a data frame. Here we compute the mean of sepal lengths.
mean(iris$Sepal.Length)


# Create a summary of Petal.Width




```

*** =solution
```{r}

# Functions take objects as arguments inside the parenthesis. Take a peek of iris.
head(iris)

# You can use the $-sign to access columns. Here you can calculcate the mean of sepal lengths.
mean(iris$Sepal.Length)

# Create a summary of Petal.Width
summary(iris$Petal.Width)



```

*** =sct
```{r}

test_function("head",args="x")
test_function("mean",args="x")
test_function("summary", args="object", 
              incorrect_msg="Did you create a summary of the Petal.Width column in iris, using `$` to access it?")

test_error()

success_msg("Good work!")

```


--- type:NormalExercise lang:r xp:150 skills:1
## Functions and arguments


As we just saw, in R you can operate on objects using *functions*. This is how you're actually going to do all the magic.

During this course you will operate on ready-made functions, but for clarity we will look at some of the details on how functions are built and learn how to use their arguments.

A function is an object that can be *called* by typing it's name and giving it the necessary *arguments* inside parenthesis.


*** =instructions
- Execute the function calls that are defined.
- Call the `do_maths()` function with `n` equal to `20` and `k` equal to `10`.
- Create object 'ans' by assigning the requested value of `do_maths()`to it.

*** =hint
- Assignment is done with the `<-` operator.
- The answer is of the form: `object <- function(arguments)`

*** =pre_exercise_code
```{r}
# code to preload datasets etc

add_5 <- function(n) {
  n + 5
}
do_maths <- function(n, k = 5) {
n * k - n + k / k
  }
```


*** =sample_code
```{r}

# Below are the definitions of two functions: add_5() and do_maths(). The goal of this exercise is to understand how to use them.

# add_5() is a simple function with one argument: n. add_5() adds 5 to n when it is called and returns the result.
add_5 <- function(n) {
  n + 5
}
# Functions can be more complicated and have multiple arguments. Arguments can also have default values.
do_maths <- function(n, k = 5) {
  n * k - n + k / k
}


# Here we call the add_5 function twice. ( add_5 is already saved in memory )
add_5(n = 5)
add_5(-5)


# Here we call the do_maths function twice.
do_maths(n = 10, k = 5)
do_maths(n = 10) # k = 5 by default

# If you give the arguments in the right order, you don't have to state which is which.
do_maths(8, 3)
do_maths(k = 3, n = 8)


# Call do_maths() and create the object 'ans' here.




```

*** =solution
```{r}

# Call do maths and create the object 'ans' here.
do_maths(20, 10)

ans <- do_maths(20, 10)



```

*** =sct
```{r}

# tests

test_function("do_maths",args=c("n","k"))
test_object("ans")

test_error()

success_msg("Great job!")

```

<<<<<<< HEAD
--- type:NormalExercise lang:r xp:150 skills:1,3
## Vectors
=======
--- type:NormalExercise lang:r xp:200 skills:1,3
## Vectors and more data frames
>>>>>>> origin/master

You can combine values of the same basic type to a *vector*, using the function `c()`. 

A vector is an object containing multiple elements. Vectors are quite important in R. The columns of a data frame (such as `iris`) are vectors.

In this execrice you will learn how to create your own vector and use it as an argument in a function.

*** =instructions
- See and execute the examples on how to create a vector
- Use any available resource to find the connection between percentiles and quantiles
- Define a vector `p` that includes the quantile probabilities corresponding to the 25th, 50th, and 75th percentiles.
- Use `?quantile` to see the help page on `quantile` and look at the arguments of the default method.
- Compute the 25th, 50th and 75th percentiles of the Petal.Length column in `iris` using `quantile()`.

*** =hint
- Remember that `$` can be used to access a column of a data frame
- Use `c()` to define `p`. It should have 3 elements.
- You then need two give 2 arguments to `quantile()`: `x` and `probs`.


*** =pre_exercise_code
```{r}
```


*** =sample_code
```{r}

# Here we create a vector with numeric elements
c(2, 3, 4, 5)

# Another one
c(0.1, 0.2, 5.84, 0.7)

# This time we'll save it and then compute it's range by calling range()
my_vector <- c(0.1, 0.2, 5.84, 0.7)
range(my_vector)


# Define a vector p of quantile probabilities corresponding to the 25th, 50th and 75th percentiles



<<<<<<< HEAD
# Get the 25th, 50th and 75th percentile of Sepal.Width in iris, using quantile(). Use the vector p as one of the arguments of quantile().
=======
# Get the 25th, 50th and 75th percentile of Sepal.Width in iris, using quantile(). Use the vector you created above as one of the arguments in quantile().
>>>>>>> origin/master
?quantile


```

*** =solution
```{r}

# Define a vector p of quantile probabilities corrsponding to the 25th, 50th and 75th percentiles
p <- c(0.25, 0.5, 0.75)

# Get the 25th, 50th and 75th percentile of Sepal.Width
quantile(iris$Petal.Length, probs = p)



```

*** =sct
```{r}

test_object("p", incorrect_msg="The vector p should include 3 elements corresponding to the 25th, 50th and 75th percentiles.")

test_function("quantile", args=c("x","probs"),
              incorrect_answer = "Did you call `quantile()` with two arguments?")

test_error()

# Final message the student will see upon completing the exercise
success_msg("Great work, that was a tough one!")

```

--- type:NormalExercise lang:r xp:100 skills:1, 3
## Generic functions

Some functions in R are very generic. They can take any object as an argument and they seem to magically do exactly what you'd hope they do in all situations.

One such example is the `summary()` function.


*** =instructions
- Execute the code that creates a summary of the `Species` in `iris`
- Then create a summary of `iris`
- `summary()` behaves differently because the *class* of `Species` and `iris` is different.

*** =hint
- Get the summary of an object by typing summary(object)

*** =pre_exercise_code
```{r}
# code to preload datasets etc
```


*** =sample_code
```{r}

# Create a summary of Species in iris
summary(iris$Species)

# Create a summary of iris



```

*** =solution
```{r}

# Create a summary of Species in iris
summary(iris$Species)

# Create a summary of iris
summary(iris)


```

*** =sct
```{r}

# tests. 


test_student_typed("summary(iris)", incorrect_msg="Please type in the code to create a summary of `iris`")

test_error()

success_msg("Great job! Next we'll take a brief look at classes.")

```

--- type:MultipleChoiceExercise lang:r xp:100 skills:1, 3
## What's your class?

Earlier we saw that `summary()` knows to give different output when it's arguments belong to different *classes*.

A good way to get information about an object is the `str()` function. It prints out the given objects class and structure.

Inspect the structure of `iris` by using `str()`. Type your code directly to the R console and execute the code by pressing `Enter`.

What is the class of `Species` in `iris`?


*** =instructions
- string
- numeric
- factor
- character

*** =hint
- Use str(iris) or str(iris$Species)

*** =pre_exercise_code
```{r}
# code to preload datasets etc
```


*** =sct
```{r}

# tests.

test_error()

msg1 = "Sorry, string is not a class in R."
msg2 = "Numeric is a class in R, but is it not correct here, even though Species has numbers as labels."
msg3 = "Yes! it is indeed a factor!"
msg4 = "Character is a class in R, but it is not correct here."
test_mc(correct=3, feedback_msgs=c(msg1, msg2, msg3, msg4))

# Final message the student will see upon completing the exercise
success_msg("Great job! Now you know the basics of R :) ")

```
