---
title       : Meet R
description : Soft introduction to R

--- type:NormalExercise lang:r xp:50 skills:1

## Basic tools

Welcome. On your right you see the R editor area and below that the R console.  

The way to use these two is to first write commands to the editor area (just as you would write in any text editor). 
Then you can execute the row where your cursor is, or the rows you have highlighted by pressing Ctrl+Enter.  

Output will then appear in the console. 

Let's try it!
  
*** =instructions
- Follow the instructions in the sript and use R as a calculator
- `+`, `-`, `/`, `*`  are the plus, minus, division and multiplication operations.
- When you are done experimenting, press the 'Submit answer' button.

*** =hint
- Remember that Ctrl+Enter executes a single row.


*** =pre_exercise_code
```{r}
```


*** =sample_code
```{r}
# I am a comment telling you that this is the R editor!


# You can use R as a calculator. Execute the below line with Ctrl + Enter
5 + 5


# Executing a comment will only print the comment, try it!
# 5 + 5


# Try your own calculations below, and execute them.



# Below me is the R console, where you will see output

```

*** =solution
```{r}
# I am a comment telling you that this is the R editor!


# You can use R as a calculator. Execute the below line with Ctrl + Enter
5 + 5


# Executing a comment will only print the comment, try it!
# 5 + 5


# Try your own calculations below, and execute them.



# Below me is the R console, where you will see output

```

*** =sct
```{r}

# It's always smart to include the following line of code at the end of your SCTs
# It will check whether executing the student's code resulted in an error, 
# and if so, will cause the exercise to fail
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```


--- type:NormalExercise lang:r xp:100 skills:1
## R objects

In R you can create and operate on things called *objects*. 
An object is just something that knows what kind of object it is and can store information.

The beauty of R is that you can store these objects for later use in your very own variables using the assign operator: `<-`.


Let's try it!

*** =instructions
- Execute the lines that define the variables `myHello` and `ans` and print their values
- Create your own variable, called `myNname` and assign your name or nickname to it.
- Create your own variable, called `myBD` and assign your birth decade to it (it's ok to lie)
- Create your own variable, called `myMessage` and assign the text "you are awsome!" to it.

*** =hint
- Remember that Ctrl+Enter executes a single row.

*** =pre_exercise_code
```{r}
```


*** =sample_code
```{r}

# create a variable named myHello
myHello <- "Hi there!"

# print the variable by simply typing it's name
myHello

# Create a numeric variable
ans <- 5 + 5

# Do some operations with ans
(ans/5) + 2


# Create myName, myBD and myMessage here



```

*** =solution
```{r}

# create a variable named myHello
myHello <- "Hi there!"

# print the variable by simply typing it's name
myHello

# Create a numeric variable
ans <- 5 + 5

# Do some operations with ans
(ans/5) + 2


# Create myName and myMessage here
myName <- "Tuomo"
myBD <- 80
myMessage <- "you are awsome!"


```

*** =sct
```{r}

test_object("myMessage")

# It's always smart to include the following line of code at the end of your SCTs
# It will check whether executing the student's code resulted in an error, 
# and if so, will cause the exercise to fail
test_error()

# Final message the student will see upon completing the exercise
success_msg(paste0(myName,"_",myBD,", ", myMessage))

```


--- type:NormalExercise lang:r xp:100 skills:1
## Title


Now that we know how to create variables that store our objects using the assign operator `<-`, the real fun can begin!

In R you can operate on objects using *functions*. This is how we're actually going to do all the magic. During this course we will ony operate on ready-made functions, but for clarity we will look at how a function is created and what are it's parts.

A function is called by typing it's name and giving it the necessary *arguments* inside parenthesis.


*** =instructions
- Try the function calls that are already defined
- use the `create_nick()`- function to create a nick with name "R" and number 93.
- assign the created nick to a variable called 'ans'

*** =hint
- Remember that the editor code is executed line by line with Ctrl + Enter
- Assignment is done with the `<-` operator: `variable_name <- function_name(arguments)

*** =pre_exercise_code
```{r}
# code to preload datasets etc
add_5 <- function(n) {
  result <- n + 5
  result
}

create_nick <- function(name = "Juha", no = 86) {
  
  paste("Your nick is: ", name, no, sep="")

}
```


*** =sample_code
```{r}

# create a simple function with one argument: n
add_5 <- function(n) {
  result <- n + 5
  result
}

# call the add_5 function. (It is already defined and saved for you)
add_5(n = 5)

# functions can be more complicated and have multiple arguments. Arguments can also have default values
create_nick <- function(name, no = 86) {
  
  paste("Your nick is: ", name, no, sep="")

}

# call the create_nick function
create_nick(name="Juha")

# If you give the arguments in the right order, you don't have to state which is which.
create_nick("Juha")
create_nick("Minni", 98)
create_nick(no = 98, name = "Minni")


# assign the function output to a variable 'ans' here





```

*** =solution
```{r}

# call the add_5() function. (It is already defined and saved for you)
add_5(n = 5)

# call the crea_nick() function
create_nick(name="Juha")

# If you give the arguments in the right order, you don't have to state which is which.
create_nick("Juha")
create_nick("Joni", 92)
create_nick(no = 92, name = "Joni")

# assign the function output to a variable 'ans' here
ans <- create_nick("R", 93)


```

*** =sct
```{r}

# tests. 

# test_object("name") 

# test_function("name", args=c("arg1","arg2"))

test_object("ans")

# It's always smart to include the following line of code at the end of your SCTs
# It will check whether executing the student's code resulted in an error, 
# and if so, will cause the exercise to fail
test_error()

# Final message the student will see upon completing the exercise
success_msg("Great job!")

```


--- type:NormalExercise lang:r xp:100 skills:1
## R data frames

The most common object to operate on is the `data frame` object, which stores data in tabular form.

Here we will look at some of the functions that are helpful for inspecting a data frame: `head()`, `tail()`, `str()`, `names()`. We will use an R data set called `iris`, which contains information about different flower species.



*** =instructions
- 

*** =hint
- Remember that Ctrl+Enter executes a single row.

*** =pre_exercise_code
```{r}
```


*** =sample_code
```{r}

# sample code



```

*** =solution
```{r}

# solution


```

*** =sct
```{r}


# It's always smart to include the following line of code at the end of your SCTs
# It will check whether executing the student's code resulted in an error, 
# and if so, will cause the exercise to fail
test_error()

# Final message the student will see upon completing the exercise
success_msg(paste0(myName,"_",myBD,", ", myMessage))

```
